|

### 2020-03-15 analysis/paper/manuscript.Rmd
Address Greg and Keith's suggestions

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples). While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore data and models in ways not amenable to static formats and, in the process, these heuristic improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In both cases interactive visuals are nested inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)). See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in relative abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988]. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. These abilities are illustrated  by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

@@

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)). RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results. The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data.

RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis. The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs. Basic inputs are those like species, season, and survey year(s). Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on. This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough the data checking and analysis required for formal stock assessment begins.

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019]. We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document. We also used the crosstalk package to link the data displayed across multiple plots. Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data. The intent of this tool was not to replace detailed documentation on the model [i.e. @cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations. For instance, explorations of the residual plots in the dashboard allowed the authors to quickly identify conflicts between two surveys used in the model. Potential solutions to this problem were then explored by modifying the configuration of the model and these changes were quickly evaluated using a dashboard that is produced in seconds.

The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file. This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model. For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model. These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables. Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieve concerns that something is hidden.

@@

The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations. Consequently, there exists the possibility of generalizing these tools to simplify their application to other regions. Region specific data management and modelling approaches, however, will surely complicate matters as tools built for one region may be difficult to apply in another region. Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications to be applied to another model being developed. Nevertheless, considerable chunks of code could be recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard. This effort was one step forward in improving the generality of the tool and another step, which is actively being worked on by a colleague in IFREMER, is to modify the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014]. Though more time is required to make the dashboard code more general as the developer must ensure the code is robust to a wider range of scenarios, substantial value could be added by making it more widely applicable. Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g. StoX software; @johnsen2019]. These are, of course, early days in the development and application of interactive tools to stock assessment. Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards. Regardless of the approach, one should strive to make the path taken transparent and reproducible.

>>>>

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools available that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples). Such tools complement static documentation as they allow audiences to explore data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts. These will be helpful tools to utilize as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid in the peer review of stock assessments.

We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In both cases interactive visuals are nested inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table \@ref(tab:pkg). See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in relative abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and, as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988]. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. These abilities are illustrated  by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

@@

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from RStrap analyses are very large and, depending on the species, may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for fisheries scientists to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)). RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results. The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data.

RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis. The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs. Basic inputs are those like species, season, and survey year(s). Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on. This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough data checking and analysis required for formal stock assessment begins.

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019]. We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document. We also used the crosstalk package to link the data displayed across multiple plots. Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data. The intent of this tool was not to replace detailed documentation on the model [i.e. @cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the stock and model. Explanatory text and detail were therefore kept to a minimum, allowing the audience to explore the figures and data unencumbered by verbose descriptions.

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations. For instance, explorations of the residual plots in the dashboard allowed the authors to quickly identify conflicts between two surveys used in the model. Potential solutions to this problem were then explored by modifying the configuration of the model and these changes were quickly evaluated using a dashboard that is produced in seconds.

The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file. This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model. For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model. These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables. Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieves any potential concerns that something is hidden.

@@

The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations but the solutions presented are specific to the Northwest Atlantic Fisheries Centre. The concept is therefore broadly applicable, however, the underlying code will require modifications to accommodate region and stock specific data management and modelling approaches. Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications when applied to a model developed for another cod stock. Nevertheless, considerable chunks of code were recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard. Additionally, a French colleague is currently modifying the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014]. Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g. StoX software; @johnsen2019]. These are, of course, early days in the development and application of interactive tools to stock assessment. Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards depending on user needs and preferences. Regardless of the approach, a common goal is to make the workflow more efficient, reproducible and transparent.
